---
title: "Homework 2"
author: "Emily Buzzelli"
date: "August 20, 2018"
output: md_document
---
##Problem 1 - Flights at ABIA
####I have family that lives in Cleveland, in Houston and in Nashville.  I am regularly taking trips to see them.  I want to use the Austin airport data to help me plan my trip.  First, I'd like to look at which carriers fly directly to these locations, and the number of flights to each location per year:
```{r echo=FALSE,results='hide',message=FALSE, warning=FALSE}

library(ggmap)
library(dplyr)

#Read in data
flights <- read.csv("ABIA.csv")
airports<- read.csv("airports_lat_long.csv")

#rename lat/long location_ID column to be "Dest"
colnames(airports)[1] <- "Dest"

#get Dest lat/Long
dplr <- left_join(flights, airports, by="Dest")

#rename last columns of new df
colnames(dplr)[31] <- "Dest_Long"
colnames(dplr)[30]<-"Dest_Lat"

#rename airport column again
colnames(airports)[1] <- "Origin"

#get Origin lat/long
dplr <- left_join(dplr, airports, by="Origin")

#rename new last two columns of df
colnames(dplr)[32] <- "Origin_Lat"
colnames(dplr)[33]<-"Origin_Long"

#add a column for total delay time
total_delay = dplr$ArrDelay+dplr$DepDelay
dplr = cbind(dplr, total_delay)


#Question to answer when should I take weekend trips to visit my family? -- parents in Ohio #(Akron or Cleveland), sister in Detroit, brother in Nashville, husband in Houston.  I want to #avoid delayes/cancellations

#create some new data frames
cleveland = dplr[dplr$Dest == "CLE",]
cleveland = rbind(cleveland, dplr[dplr$Origin == "CLE",])
text = rep("cleveland", length(cleveland$Origin))
cleveland = cbind(cleveland, text)


nashville = dplr[dplr$Dest == "BNA",]
nashville = rbind(nashville, dplr[dplr$Origin == "BNA",])
text = rep("nashville", length(nashville$Origin))
nashville = cbind(nashville, text)


houston = dplr[dplr$Dest == "HOU",]
houston = rbind(houston, dplr[dplr$Origin == "HOU",])
text = rep("houston", length(houston$Origin))
houston = cbind(houston, text)


#See which carriers operate the flights I care about
familylocations = rbind(cleveland, nashville, houston)
counts = table(familylocations$UniqueCarrier, familylocations$text)
counts = counts[rowSums(counts)>0,]
barplot(counts, main="Number of flights", xlab="Location",col=c("darkblue","red"), legend = rownames(counts))




```

####Looks like I don't have that many choices of carriers if I want direct flights to these locations.  It also looks like the largest number of flights exist between Austin and Houston.  Now I want to use this data to understand when I should plan trips to these 3 cities to avoid delays.  First let's take a look at the total delay time for flights going to or returning from each of these cities by month.  This will help me understand what time of year I should plan my trips:


```{r echo=FALSE,results='hide',message=FALSE, fig.width = 11, fig.height=7}
#let's look at everything on one plot

#plot with invisible boxes
boxplot(familylocations$total_delay~as.factor(familylocations$Month),main = "Total Delay Time for each Location by Month", xlab = "Month", ylab = "Total Delay in minutes", xlim = c(.25, 12.75), ylim=c(-30,800 ), boxfill=rgb(1, 1, 1, alpha=1), border=rgb(1, 1, 1, alpha=1))

#plot houston delays
boxplot(familylocations[familylocations$text=="houston",]$total_delay~ as.factor(familylocations[familylocations$text=="houston",]$Month), xaxt = "n", add = TRUE, boxfill="red", boxwex=0.25, at = 1:12 - 0.25) #shift these left by -0.25

#plot nashville delays
boxplot(familylocations[familylocations$text=="nashville",]$total_delay~ as.factor(familylocations[familylocations$text=="nashville",]$Month), xaxt = "n", add = TRUE, boxfill="blue", boxwex=0.25, at = 1:12 + 0.25) #shift these right by .25

#plot cleveland delays
boxplot(familylocations[familylocations$text=="cleveland",]$total_delay~ as.factor(familylocations[familylocations$text=="cleveland",]$Month), xaxt = "n", add = TRUE, boxfill="green", boxwex=0.25, at = 1:9 + 0.0) #shift these left by 0

#add legend
legend("topleft", legend=c("Houston", "Cleveland", "Nashville"), col=c("red","green", "blue"), lwd = 2, cex=0.8)

#add a grid for easier reading
grid(nx = NULL, ny = NULL, col = "lightgray", lty = "dotted", lwd = par("lwd"), equilogs = TRUE)



```

####From this plot, I can see that the best time to travel to Houston is probably sometime in the fall if I want to avoid delays (sept, oct and november) seem to have the least variability and most common occurances around 0 minutes for delay.  For travelling to Nashville I can see that January, September and November would be the optimal time for a trip.  In these months, it looks like flights tend to leave a bit early and do not have as many delays as other months.  Unfortunately, it looks like direct flights to Cleveland ceased in September of 2008, so I might not be able to get a direct flight here anymore, but when they were occuring, it looks like the best months to travel would have been over the summer (specifically July and August) if I wanted the shortest delay.

####Now let's take a look at the days of the week with the shortest delays for flights TO each city from Austin and flights FROM each city from Austin.  This should tell me what days to leave and what days to return in order to avoid delays.  First, let's look at delays of flights leaving Austin by destination location/day of the week.
```{r echo=FALSE,results='hide',message=FALSE,fig.width = 11, fig.height=7}
#let's look at everything on one plot
#plot with invisible boxes
boxplot(familylocations$total_delay~as.factor(familylocations$DayOfWeek),main = "Total Delay Time by Day for Flights Departing From Austin", xlab = "Day of Week", ylab = "Total Delay in minutes", xlim = c(.25, 7.75), ylim=c(-30,800 ), boxfill=rgb(1, 1, 1, alpha=1), border=rgb(1, 1, 1, alpha=1))

#plot delays on flights arriving from houston
boxplot(houston[houston$Origin=="AUS",]$total_delay~ as.factor(houston[houston$Origin=="AUS",]$DayOfWeek), xaxt = "n", add = TRUE, boxfill="red", boxwex=0.25, at = 1:7 - 0.25) #shift these left by -0.25

#plot nashville delays
boxplot(nashville[nashville$Origin=="AUS",]$total_delay~ as.factor(nashville[nashville$Origin=="AUS",]$DayOfWeek), xaxt = "n", add = TRUE, boxfill="blue", boxwex=0.25, at = 1:7 + 0.25) #shift these right by .25

#plot cleveland delays
boxplot(cleveland[cleveland$Origin=="AUS",]$total_delay~ as.factor(cleveland[cleveland$Origin=="AUS",]$DayOfWeek), xaxt = "n", add = TRUE, boxfill="green", boxwex=0.25, at = 1:7 + 0.0) #shift these left by 0

#add legend
legend("topleft", legend=c("Houston", "Cleveland", "Nashville"), col=c("red","green", "blue"), lwd = 2, cex=0.8)

#add a grid for easier reading
grid(nx = NULL, ny = NULL, col = "lightgray", lty = "dotted", lwd = par("lwd"), equilogs = TRUE)



```

####From the plot above I can see that the best days to leave Austin for Cleveland are Sunday (day 7) and Monday (day 1), which tend to have flights that leave early.  The best day to leave Austin for Houston is Saturday.  The best day to leave Austin for Nashville looks to be Saturday or Tuesday - but most of the days seem flights are generally on time.  In fact, this seems to be the case for most of the flights departing Austin for these cities.  Let's see how flights arriving in Austin from these cities look like.  This should give some details on what days of the week I should return to Austin to avoid delays.

```{r echo=FALSE,results='hide',message=FALSE,fig.width = 11, fig.height=7}
#let's look at everything on one plot
#plot with invisible boxes
boxplot(familylocations$total_delay~as.factor(familylocations$DayOfWeek),main = "Total Delay Time by Day for Flights Arriving in Austin", xlab = "Day of Week", ylab = "Total Delay in minutes", xlim = c(.25, 7.75), ylim=c(-30,800 ), boxfill=rgb(1, 1, 1, alpha=1), border=rgb(1, 1, 1, alpha=1))

#plot delays on flights arriving from houston
boxplot(houston[houston$Dest=="AUS",]$total_delay~ as.factor(houston[houston$Dest=="AUS",]$DayOfWeek), xaxt = "n", add = TRUE, boxfill="red", boxwex=0.25, at = 1:7 - 0.25) #shift these left by -0.25

#plot nashville delays
boxplot(nashville[nashville$Dest=="AUS",]$total_delay~ as.factor(nashville[nashville$Dest=="AUS",]$DayOfWeek), xaxt = "n", add = TRUE, boxfill="blue", boxwex=0.25, at = 1:7 + 0.25) #shift these right by .25

#plot cleveland delays
boxplot(cleveland[cleveland$Dest=="AUS",]$total_delay~ as.factor(cleveland[cleveland$Dest=="AUS",]$DayOfWeek), xaxt = "n", add = TRUE, boxfill="green", boxwex=0.25, at = 1:7 + 0.0) #shift these left by 0

#add legend
legend("topleft", legend=c("Houston", "Cleveland", "Nashville"), col=c("red","green", "blue"), lwd = 2, cex=0.8)

#add a grid for easier reading
grid(nx = NULL, ny = NULL, col = "lightgray", lty = "dotted", lwd = par("lwd"), equilogs = TRUE)


```

####From the plot above, it seems like the best days to arrive in Austin from Cleveland are either Monday (day 1) or Saturday (day 6).  Also from the plot I can see that the day to avoid arriving in to Austin from Cleveland is Sunday, which seems to have a larger number of delays on average.  The best days to arrive from Houston look to be Monday, Tuesday or Saturday.  The best days to arrive from Nashville look to be Monday or Saturday and similarly to Cleveland, it seems that arriving from Nashville on Sunday tends to have larger delay times.



##Problem 2 - Author Attribution
####First I pulled all of the text out of the list of documents & created a corpus for the training data
```{r echo=FALSE,results='hide',message=FALSE}
library(tm)

#Wrapper function
readerPlain = function(fname){
  readPlain(elem=list(content=readLines(fname)), 
            id=fname, language='en') }



author_dirs_train = Sys.glob('~/ReutersC50/C50train/*')
author_dirs_test = Sys.glob('~/ReutersC50/C50test/*')

file_list_train = NULL
labels = NULL
for(author in author_dirs_train) 
{
  author_name = substring(author, first = 29)
  files_to_add = Sys.glob(paste0(author, '/*.txt'))
  file_list_train = append(file_list_train, files_to_add)
  labels = append(labels, rep(author_name, length(files_to_add)))
}

#Getting rid of '.txt' from filename
all_docs_train = lapply(file_list_train, readerPlain) 
names(all_docs_train) = file_list_train
names(all_docs_train) = sub('.txt', '', names(all_docs_train))



file_list_test = NULL
labels = NULL
for(author in author_dirs_test) 
{
  author_name = substring(author, first = 29)
  files_to_add = Sys.glob(paste0(author, '/*.txt'))
  file_list_test = append(file_list_test, files_to_add)
  labels = append(labels, rep(author_name, length(files_to_add)))
}

#Getting rid of '.txt' from filename
all_docs_test = lapply(file_list_test, readerPlain) 
names(all_docs_test) = file_list_test
names(all_docs_test) = sub('.txt', '', names(all_docs_test))


my_corpus = Corpus(VectorSource(all_docs_train))
test_corpus = Corpus(VectorSource(all_docs_test))

my_corpus = tm_map(my_corpus, content_transformer(tolower)) # make everything lowercase
my_corpus = tm_map(my_corpus, content_transformer(removeNumbers)) # remove numbers
my_corpus = tm_map(my_corpus, content_transformer(removePunctuation)) # remove punctuation
my_corpus = tm_map(my_corpus, content_transformer(stripWhitespace)) ## remove excess white-space
my_corpus = tm_map(my_corpus, content_transformer(removeWords), stopwords("SMART"))

```


##Problem 3 - Practice with Association Rule Mining

####For this problem, let's first plot the rules output of the apiori algorithm based on lift, support & confidence 

```{r echo=FALSE,results='hide',message=FALSE, warning = FALSE, fig.width = 11, fig.height=7}
library(tidyverse)
library(arules)  
library(arulesViz)


# Read in baskets
baskets = readLines("groceries.txt")
baskets <- strsplit(baskets, ",")


#change to transactions
baskettrans = as(baskets, "transactions")


# Now run the 'apriori' algorithm
# Look at rules with support > .005 & confidence >.1 & length (# artists) <= 5
basketrules = apriori(baskettrans, parameter=list(support=.005, confidence=.1, maxlen=5))

#plot
plot(basketrules)

```


####From this plot, let's pick a range of variables that will give us rules with decent support, confidence & lift.  Let's try support<.015, confidence > .35 and lift >2.5

```{r echo=FALSE,results='hide',message=FALSE, warning = FALSE, fig.width = 11, fig.height=7}
sub1 = subset(basketrules, subset=confidence > 0.35 & support < 0.015 & lift> 2.5)
summary(sub1)
plot(sub1, method='graph', zoom = 4)

```

####From this plot, we can see some clear associations between dairy products - such as butter, sour cream, yogurt & milk.  We also see clusterings of fruit and vegetables, meaning purchases of one kind of these items are associated to other kinds of these items.  We can also see that there are some items such as whole milk, vegetables & fruit are common in all baskets.
